---
title: "Assignment Task"
author: "Kwanele Fadane (202000003)"
date: "17/11/2020"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

# Basics
## 1.1 Variables
What is the value of " B " after running these command?
```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t
```

The value of B is: `r round(B,3)`

What type of vector do you get in each of the following?
```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```

For the 1st vector we get : `r typeof(vec_1)`

For the 2nd Vector we get : `r typeof(vec_2)`

For the 3rd Vector we get : `r typeof(vec_3)`

For the 4th Vector we get : `r typeof(vec_4)`

### 1.2 Conditional Subsetting
Consider the following vector:
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
```
### 1.2.1 Create a vector vec_1 with no missing values using conditional subsetting and explain your code
```{r}
vec_1 <- na.omit(vec)
```

`r vec_1`

The `na.omit()` function lets us remove all incomplete objects in a data set. In the code the 'NA' are the missing values and the code will remove every 'NA'.

### 1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code 

```{r}
vec_2 <- vec[c(FALSE, FALSE, FALSE, TRUE, FALSE, FALSE,FALSE, TRUE ,FALSE, FALSE, FALSE, FALSE, TRUE)]
```

`r vec_2`

The logic vector lets us use the `TRUE` and `FALSE` to identify the complete and incomplete variable. In this code `FALSE` is allocated to the complete variable and the `TRUE` to the incomplete. Therefore `TRUE` will select all NA variables.
 
 
### 1.2.3 Write a code to determine how many missing values are in vec.

```{r}
vec_3 <- vec[c(FALSE, FALSE, FALSE, TRUE, FALSE, FALSE,FALSE, TRUE ,FALSE, FALSE, FALSE, FALSE, TRUE)]
```

`r length(vec_3)`

### 1.2.4. Write a R code to create a data frame which contains details of 3 students as follows
### Name
### Surname
### Age
### Gender

```{r}
students = data.frame(Name=c("Andiswa", "Athenkosi", "Yanga"),
                     Surname=c("Van Vyk", "Bossman", "Buthelezi"),
                     Age=c(20,22,25),
                     Gender=c("Male", "Female", "Male")
                     )
print(students)
```


### 2. Functions in R
### 2.1. Create a function which returns a mean of an input vector and explain your code
```{r}
vec_one <- c(2,4,8, NA, 9, 12, NA, 16)
vec_two <- c(2,5,NA,7,9)

mean_value <-function(input){
  output <-mean(input, na.rm = TRUE)
  return(output)
}
```

Mean value for vec_one is : `r mean_value(vec_one)`

Mean value for vec_two is : `r mean_value(vec_two)`


Assign `mean_value` to a function with an input parameter which gives the mean value of a given vector. The function will have take one argument which are the vectors, `vec_one ` and `vec_two`. The `output` variable is assigned to a `mean` function.The `mean` function looks at the values of the vectors and `na.rm = TRUE` removes the missing values from the input vector.

### 2.2. Create a function which take no inputs , it reads a Titanic .csv file and returns its content and explain your code

```{r}
library(tidyverse)
df<-tibble(read.csv("train.csv"))
print(df)

```

The library `tidyverse` is for using read.csv. We assign the variable df to read.csv which is used to gives us accesses to the `train` document, then we print the df variable to get all the information on the file.

### 3. Conditionals and Loops

### 3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse

```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE")
```

```{r}
gender <-function(input){
  output <-ifelse(input == "MALE",1,2)
  return(output)
}
```

`r gender(g)`

### 3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
```

### 3.2.1. Use the "cbind " function in R to create one dataset and assign it to df.

```{r}
df <- tibble(cbind(x,y))
print(df)
```

### 3.2.2. Convert df from 3.2.1 into a data.frame

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)

df1 <-tibble(data.frame(df))
print(df1)
```

### 3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F") in 3.2.1

```{r}
library(tidyverse)
df %>%  group_by(cbind(y)) %>%  count
```

